Probability Aggregation: Dynamic Hierarchical Modeling of Sparse Expert Beliefs

### 2013-08-31 paper.tex
Edits

\begin{abstract}
Most subjective probability aggregation procedures use a single probability judgement from each expert, even though it is common for experts studying real problems to update their probability estimates over time.  This paper makes an advance towards unexplored areas of probability aggregation by considering a dynamic context in which experts are allowed to update their beliefs at random intervals. The updates are assumed to occur very infrequently, resulting into a highly sparse dataset that cannot be modeled by standard time-series procedures. In response to the lack of appropriate methodology, this paper presents a dynamic model that takes into account the forecaster's level of self-reported expertise and produces aggregate probabilities that are sharp and well-calibrated both in- and out-of-sample. The model is demonstrated on a real-world dataset that includes over 2,300 experts making multiple probability forecasts on different subsets of 166 different international political events. 
\end{abstract}

@@

\section{Introduction}
\label{intro}
Individual experts can differ radically from one another in their abilities to assess probabilities of future events. The quality of a probability assessment is typically evaluated in regards to \textit{calibration}, which is an attribute measuring the degree to which the frequency of occurrence of the events agree with the given probabilities. This means that, for instance, the proportion of occurred events is 60\% for all those events for which a well-calibrated expert assessed a probability of 0.60. Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), relative differences can occur between different types of experts. In particular, \citet{wright1994coherence} argue that a higher level of self-reported expertise on the subject-matter can be associated with better calibration.

Calibration by itself, however, is not sufficient for a useful probability assessment. To see this, consider a relatively stationary process, such rain on different days in some geographic region, whose empirical frequency of occurrence in the last 10 years is 45\%. In this setting an expert can assess a constant probability of 0.45 and be well-calibrated. This assessment, however, can be made without any subject-matter expertise or special training. For this reason the long-term frequency is often considered as the baseline probability -- a naive assessment that provides the decision-maker very little extra information. A consulting expert should therefore aim to make probability assessments that are as far from the baseline as possible. The extent to which his probabilities differ from the baseline is measured with an important attribute called \textit{sharpness} (\citet{gneiting2008rejoinder, winkler2008comments}). If the expert is both sharp and well-calibrated, he is able to forecast the behavior of the process with high certainty and accuracy. From the decision-maker's perspective, his advice is highly useful as it involves much information and very little risk. Therefore useful probability estimation should aim to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, murphy1987general}). This is a well-defined goal that has led to a wide range of novel and insightful observations in probability forecasting.

@@

The paper begins with a description of our probability forecasting data that were collected by asking over 2,300 experts to give probability forecasts and to self-assess their level of expertise on a subset of 166 geopolitical binary events. After summarizing the dataset, the paper introduces a dynamic hierarchical model for capturing the crowd belief. The model is estimated in a two-step procedure. First, a sampling step produces constrained parameter estimates via Gibbs sampling (see \citet{geman1984stochastic} for the original introduction of Gibbs sampling). Second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure. An extension of this model to polychotomous outcomes is briefly discussed before model evaluation. The first evaluation section uses synthetic data to study the accuracy to which the two-step procedure is able to estimate parameter values. The second evaluation section discusses different aspects of the model by applying it to our real-world probability forecasting data. The paper concludes with a discussion on future research directions and model limitations.

@@

The data collection began with a recruitment of 2,365 forecasters ranging from graduate students to political science faculty and practitioners. The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth. Requirements included at least a Bachelor's degree and completion of psychological and political tests that took roughly two hours. These measures assessed cognitive styles, cognitive abilities, personality traits, political attitudes, and real-world knowledge. The experts were asked to give probability forecasts (to the second decimal point) and to self-assess their level of expertise (on a 1-to-5 scale with 1 = Not At All Expert and 5 = Extremely Expert) on a number of 166 geopolitical binary events taking place between September 29, 2011 and May 8, 2013. Each question was active for a period of time during which the participating experts were allowed to update their forecasts as frequently as they liked without being penalized. The forecasters knew that their probability estimates would be assessed for accuracy using Brier scores\footnote{The Brier score is the squared distance between the probability forecast and the event indicator that equals 1.0 or 0.0 depending on whether the event happened or not, respectively. See \citet{Brier} for the original introduction.}. This incentivized them to report their true beliefs instead of attempting to game the system (\citet{winkler1968good}). In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the forecasters also received status rewards for their performance via leader-boards displaying Brier scores for the top 20 forecasters. Since a typical expert participated only in a small subset of the 166 questions, the forecasters are considered indistinguishable conditional on the level of self-reported expertise.
Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.  For more details on the data and its collection see \citet{ungar2012good}.

@@

\caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset. The shadings represents the self-reported expertise of the forecaster who provided the probability forecast.}
\label{ExamplePlots}
\end{figure}

To illustrate the nature of the data with some concrete examples, Figures \ref{Example1} and \ref{Example2} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}. The points have been jittered slightly to make overlaps visible. The darkness of the points is positively associated with the self-reported expertise. Given that the European bailout fund was ratified before 1 November 2011 and that the Nikkei 225 index finished trading at around 8,700 on 30 September 2011, the general trends of the probability forecasts converge to the correct answers. There is, however, much disagreement among the experts. This disagreement persists even near the closing dates of the questions. Unfortunately, the data is not only noisy but also quite sparse. As mentioned before, the forecasters were allowed to update their forecast as long as the questions remained open. Updating, however, was done on a very infrequent basis. An expert gave, on average, only 0.016 forecasts per day, and so a typical expert made a forecast about every 62 days. As a result the average response rate was around 13.5 forecasts per day from a large group of experts.

\section{Model}
\label{model}
Let $p_{i,t,k} \in (0,1)$ be the probability forecast given by the $i$th expert at time $t$ for the $k$th question, where $i = 1, \dots, I_k$, $t = 1, \dots, T_k$, and $k = 1, \dots, K$. Denote the logit-probabilities with 
\begin{eqnarray*}
Y_{i,t,k} = \logit(p_{i,t,k}) = \log \left( \frac{p_{i,t,k}}{1-p_{i,t,k}}\right) \in \R
\end{eqnarray*}
and collect the logit-probability forecasts given for question $k$ at time $t$ into a vector $\boldsymbol{Y}_{t,k} = [Y_{1,t,k}\ Y_{2,t,k}\ \dots\ Y_{I_{k},t,k}]^T$. Partition the experts into $J$ groups based on some individual feature, such as self-reported expertise, with each group sharing a common multiplicative bias term $b_{j} \in \R$ for $j = 1, \dots, J$. Collect these bias terms into a bias vector $\boldsymbol{b} = [b_{1}\ b_{2}\ \dots\ b_{J}]^T$ and let $\boldsymbol{M}_k$ be a $I_k \times J$ matrix denoting  the group-memberships of the forecasters in question $k$. That is, if the $i$th expert participating in the $k$th question belongs to the $j$th group, then the $i$th row of $\boldsymbol{M}_k$ is the $j$th standard basis vector, $\boldsymbol{e}_j$. The bias vector $\boldsymbol{b}$ does not include a subindex because it is considered shared among all the $K$ questions. To secure model identifiability, it is sufficient to share only one of the elements of $\boldsymbol{b}$ among the questions. This element defines a baseline under which it is possible to estimate the remaining $J-1$ bias terms separately within each of the questions. In this paper, however, the entire vector $\boldsymbol{b}$ is shared because some of the questions in our real-world data set involve very few experts with the highest level of self-reported expertise. Under this notation, the model for the $k$th question can be expressed as 
\begin{eqnarray}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} X_{t, k} + \boldsymbol{v}_{t, k} \label{observedpr} \\
X_{t, k} &=& \gamma_k X_{t-1, k} + w_{t, k} \label{hiddenpr}\\
X_{0,k} &\sim& \mathcal{N}(\mu_0, \sigma^2_0) \nonumber
\end{eqnarray}
where Equation (\ref{observedpr}) denotes the observed process, Equation (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$. The error terms are independent and identically distributed normal random variables with mean zero
\begin{eqnarray*}
\boldsymbol{v}_{t, k} | \sigma^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}(0, \tau^2_k),
\end{eqnarray*}
and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters chosen \textit{a priori}. The hidden state $X_{t,k}$ represents the sharp and well-calibrated logit-probability for the $k$th event given the information available up to and including time $t$. To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$. These probabilities are made estimable by assuming that the observed process is associated with the filtration $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ and hence can be used as a proxy for the amount of information available at any given time point. 
%Therefore it is not unreasonable to interpreted the hidden process as the crowd belief across time.

To give intuitive interpretations of the other model parameters, notice that the error term in the observed process has mean zero. Therefore the forecasters are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from the calibrated crowd belief. An individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$. Unfortunately, due to the high sparsity of our data, estimating a bias term and analyzing the bias separately for each expert is not possible. See Section \ref{ExpertBias} for an analysis and discussion on the bias terms. Any other deviation from the calibrated crowd belief is considered \textit{random noise}. This noise is measured in terms of $\sigma^2_k$ and can be assumed to be caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions.

@@

The most challenging part of our model estimation is to incorporate the event indicators $\{Z_k\}_{k=1}^K$ such that the hidden process is well-calibrated and the interpretability of our model is not compromised. This section introduces a two-step procedure, called \textit{Sample-Then-Calibrate} or simply \textit{STC}, that achieves this goal in a flexible and efficient manner by first estimating  the model parameters under a constraint (\textit{Sampling Step}) and then performing a one-dimension optimization procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).

\subsection{Sampling Step}
\label{sampling_step}
Since $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable and the parameter estimates tend to drift during the sampling process. A well-known solution is to choose one of the elements of $\boldsymbol{b}$, say $b_3$, as the reference point and fix $b_3 = 1$.  Denote the constrained version of the model by
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
X_{t, k}(1) &=& \gamma_k(1) X_{t-1, k}(1) + w_{t, k}\\
\boldsymbol{v}_{t, k} | \sigma^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k(1) \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}\left(0, \tau^2_k(1)\right),
\end{eqnarray*}
where the trailing input parameter emphasizes the constraint $b_{3} = 1$. Since this version is identifiable, estimates of the model parameters can be obtained. Denote the estimates by placing a hat on the parameter symbol. For instance, $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ represent the estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$, respectively. These estimates are found via Gibbs sampling that only makes use of standard distributions. See Appendix \ref{appendix} for the technical details of our sampler, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.

\subsection{Calibration Step}
\label{calibration_step}
Since the model parameters can be estimated under a constrained version of $b_3$, the next step is to determine how $b_3$ should be constrained such that both the sharpness and calibration of the estimated hidden process are maximized. This section introduces an efficient procedure that finds the optimal constraint without requiring any additional runs of the sampling step. First, assume that parameter estimates  $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ have already been obtained via the constrained sampling step described in Section \ref{sampling_step}. Since for any $\beta \in \R / \{0\}$,
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
 &=&   \boldsymbol{M}_k \left(\boldsymbol{b} (1) \beta \right) \left(X_{t, k}(1)/\beta\right)  + \boldsymbol{v}_{t, k}\\
&=&   \boldsymbol{M}_k \boldsymbol{b} (\beta) X_{t, k}(\beta) + \boldsymbol{v}_{t, k},
\end{eqnarray*}
the parameter values under $b_{3} = \beta$ can be obtained from $\boldsymbol{b} (\beta) = \boldsymbol{b} (1) \beta $ and $X_{t, k}(\beta) = X_{t, k}(1)/\beta$. 
%Furthermore, an estimate of $X_{t, k}(\beta)$ is given by $\hat{X}_{t, k}(1)/\beta$. 
This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$. Since the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated by the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$. A natural criterion for this maximization is provided by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}). Due to the possibility of \textit{complete separation} in any one question (see, e.g., \citet{gelman2008weakly}), the maximization must be performed over multiple questions. Therefore we estimate
\begin{eqnarray}
 \hat{\beta}&=&  \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / \beta \right) \label{OSE}
\end{eqnarray}
where $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. The function $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
\begin{eqnarray*}
S_{BRI}(Z, X) &=& -(Z - \logit^{-1}(X))^2
\end{eqnarray*}
or the logarithmic score (\citet{good1952rational})
\begin{eqnarray*}
S_{LOG}(Z, X) &=& Z \log\left(\logit^{-1}(X)\right) + (1-Z) \log\left(1-\logit^{-1}(X)\right)
\end{eqnarray*}
Since it is not clear which rule should be used for predicting geopolitical events, such as the ones in our dataset, the \textit{Sample-Then-Calibrate} procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}. Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
\begin{eqnarray}
 \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{\beta} \nonumber\\
 \hat{\boldsymbol{b}}_{t,k}&=& \hat{\boldsymbol{b}}(1) \hat{\beta} \nonumber\\
   \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1) \hat{\beta}^2\nonumber\\
  \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\nonumber\\
  \hat{\gamma}_{k}&=& \hat{\gamma}_{k}(1)\nonumber
\end{eqnarray}
Notice that the parameters $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint. Therefore their constrained and unconstrained estimates are the same.

>>>>

\begin{abstract}
Most subjective probability aggregation procedures use a single probability judgement from each expert, even though it is common for experts studying real problems to update their probability estimates over time.  This paper makes an advance towards unexplored areas of probability aggregation by considering a dynamic context in which experts are allowed to update their beliefs at random intervals. The updates are assumed to occur very infrequently, resulting into a highly sparse dataset that cannot be modeled by standard time-series procedures. In response to the lack of appropriate methodology, this paper presents a hierarchical model that takes into account the expert's level of self-reported expertise and produces aggregate probabilities that are sharp and well-calibrated both in- and out-of-sample. The model is demonstrated on a real-world dataset that includes over 2,300 experts making multiple probability forecasts on different subsets of 166 international political events. 
\end{abstract}

@@

\section{Introduction}
\label{intro}
Individual experts can differ radically from one another in their abilities to assess probabilities of future events. Their probability assessments are often evaluated and compared in regards to \textit{calibration} that measures how closely the frequency of occurrence of the events agree with the given probabilities. For instance, the proportion of occurred events is 60\% for all those events for which a well-calibrated expert assessed a probability of 0.60. Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), relative differences can occur among different types of experts. In particular, \citet{wright1994coherence} argue that a higher level of self-reported expertise can be associated with better calibration.

Calibration by itself, however, is not sufficient for a useful probability assessment. To see this, consider a relatively stationary process, such as rain on different days in some geographic region, whose empirical frequency of occurrence in the last 10 years is 45\%. In this setting an expert can assess a constant probability of 0.45 and be well-calibrated. This assessment, however, can be made without any subject-matter expertise or special training. For this reason the long-term frequency is often considered as the baseline probability -- a naive assessment that provides the decision-maker very little extra information. Therefore a consulting expert should aim to make probability assessments that are as far from the baseline as possible. The extent to which his probabilities differ from the baseline is measured with an important attribute called \textit{sharpness} (\citet{gneiting2008rejoinder, winkler2008comments}). If the expert is both sharp and well-calibrated, he is able to forecast the behavior of the process with high certainty and accuracy. From the decision-maker's perspective his advice is highly useful as it involves much information and very little risk. Therefore, to summarize, useful probability estimation should aim to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, murphy1987general}). This is a well-defined goal that has led to a wide range of novel and insightful observations in probability forecasting.

@@

The paper begins with a description of our probability forecasting data that were collected by asking over 2,300 experts to give probability forecasts and to self-assess their level of expertise on a subset of 166 geopolitical binary events. After summarizing the dataset, the paper introduces a dynamic hierarchical model for capturing the crowd belief. The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see \citet{geman1984stochastic} for the original introduction of Gibbs sampling); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure. An extension of this model to polychotomous outcomes is briefly discussed before model evaluation. The first evaluation section uses synthetic data to study the accuracy to which the two-step procedure is able to estimate parameter values. The second evaluation section illustrates different aspects of the model by applying it to our real-world probability forecasting data. The paper concludes with a discussion on future research directions and model limitations.

@@

The data collection began with a recruitment of 2,365 experts ranging from graduate students to political science faculty and practitioners. The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth. Requirements included at least a Bachelor's degree and completion of psychological and political tests that took roughly two hours. These measures assessed cognitive styles, cognitive abilities, personality traits, political attitudes, and real-world knowledge. The experts were asked to give probability forecasts (to the second decimal point) and to self-assess their level of expertise (on a 1-to-5 scale with 1 = Not At All Expert and 5 = Extremely Expert) on a number of 166 geopolitical binary events taking place between September 29, 2011 and May 8, 2013. Each question was active for a period of time during which the participating experts were allowed to update their forecasts as frequently as they liked without being penalized. The experts knew that their probability estimates would be assessed for accuracy using Brier scores\footnote{The Brier score is the squared distance between the probability forecast and the event indicator that equals 1.0 or 0.0 depending on whether the event happened or not, respectively. See \citet{Brier} for the original introduction.}. This incentivized them to report their true beliefs instead of attempting to game the system (\citet{winkler1968good}). In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the experts received status rewards for their performance via leader-boards displaying Brier scores for the top 20 experts. Since a typical expert participated only in a small subset of the 166 questions, the experts are considered indistinguishable conditional on the level of self-reported expertise.

Overall, the data is quite sparse. As mentioned before, the experts were allowed to update their forecasts as long as the question remained open. Updating, however, was done on a very infrequent basis. An expert gave, on average, only 0.016 forecasts per day. Therefore a typical expert made a forecast about every 62 days, resulting to an average response rate of around 13.5 forecasts per day from a very large and diverse group of experts. Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.  For more details on the data and its collection see \citet{ungar2012good}.

@@

\caption[Optional caption for list of figures]{Scatterplots of the probability forecasts given for two questions in our dataset. The shadings represents the self-reported expertise of the expert who provided the probability forecast.}
\label{ExamplePlots}
\end{figure}

To illustrate the nature of the data with some concrete examples, Figures \ref{Example1} and \ref{Example2} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}. The points have been jittered slightly to make overlaps visible. The darkness of the points is positively associated with the self-reported expertise. Given that the European bailout fund was ratified before November 1, 2011 and that the Nikkei 225 index finished trading at around 8,700 on September 30, 2011, the general trend of the probability forecasts tends to converge towards the correct answers. The individual experts, however, disagree rather strongly, with the disagreement persisting even near the closing dates of the questions.

\section{Model}
\label{model}
Let $p_{i,t,k} \in (0,1)$ be the probability forecast given by the $i$th expert at time $t$ for the $k$th question, where $i = 1, \dots, I_k$, $t = 1, \dots, T_k$, and $k = 1, \dots, K$. Denote the logit-probabilities with 
\begin{eqnarray*}
Y_{i,t,k} = \logit(p_{i,t,k}) = \log \left( \frac{p_{i,t,k}}{1-p_{i,t,k}}\right) \in \R
\end{eqnarray*}
and collect the logit-probability forecasts given for question $k$ at time $t$ into a vector $\boldsymbol{Y}_{t,k} = [Y_{1,t,k}\ Y_{2,t,k}\ \dots\ Y_{I_{k},t,k}]^T$. Partition the experts into $J$ groups based on some individual feature, such as self-reported expertise, with each group sharing a common multiplicative bias term $b_{j} \in \R$ for $j = 1, \dots, J$. Collect these bias terms into a bias vector $\boldsymbol{b} = [b_{1}\ b_{2}\ \dots\ b_{J}]^T$. Let $\boldsymbol{M}_k$ be a $I_k \times J$ matrix denoting  the group-memberships of the experts in question $k$; that is, if the $i$th expert participating in the $k$th question belongs to the $j$th group, then the $i$th row of $\boldsymbol{M}_k$ is the $j$th standard basis vector $\boldsymbol{e}_j$. The bias vector $\boldsymbol{b}$ does not include a subindex because it is considered shared among all the $K$ questions. To secure model identifiability, it is sufficient to share only one of the elements of $\boldsymbol{b}$ among the questions. This element defines a baseline under which it is possible to estimate the remaining $J-1$ bias terms separately within each of the questions. In this paper, however, the entire vector $\boldsymbol{b}$ is shared because some of the questions in our real-world data set involve very few experts with the highest level of self-reported expertise. Under this notation, the model for the $k$th question can be expressed as 
\begin{eqnarray}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} X_{t, k} + \boldsymbol{v}_{t, k} \label{observedpr} \\
X_{t, k} &=& \gamma_k X_{t-1, k} + w_{t, k} \label{hiddenpr}\\
X_{0,k} &\sim& \mathcal{N}(\mu_0, \sigma^2_0) \nonumber
\end{eqnarray}
where Equation (\ref{observedpr}) denotes the observed process and Equation (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$. The error terms are independent and identically distributed normal random variables with mean zero
\begin{eqnarray*}
\boldsymbol{v}_{t, k} | \sigma^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}(0, \tau^2_k),
\end{eqnarray*}
and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters chosen \textit{a priori}. The hidden state $X_{t,k}$ represents the sharp and well-calibrated logit-probability for the $k$th event given the information available up to and including time $t$. To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$. These probabilities are made estimable by assuming that the observed process is associated with the filtration $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ and hence can be used as a proxy for the amount of information available at any given time point. 
%Therefore it is not unreasonable to interpreted the hidden process as the crowd belief across time.

To give intuitive interpretations of the other model parameters, notice that the error term in the observed process has mean zero. Therefore the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from the calibrated crowd belief. An individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$. Unfortunately, due to the high sparsity of our data, estimating a bias term and analyzing the bias separately for each expert is not possible. See Section \ref{ExpertBias} for an analysis and discussion on the bias terms. Any other deviation from the calibrated crowd belief is considered \textit{random noise}. This noise is measured in terms of $\sigma^2_k$ and can be assumed to be caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions.

@@

The main challenge is to capture a well-calibrated estimate of the hidden process without sacrificing the interpretability of our model. This section introduces a two-step procedure, called \textit{Sample-Then-Calibrate} (STC), that achieves this goal in a flexible and efficient manner: the first step estimates the model parameters under a constraint (\textit{Sampling Step}), and the second step performs a one-dimension optimization procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).

\subsection{Sampling Step}
\label{sampling_step}
Since $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable. As a result, the parameter estimates tend to drift during the sampling process. A well-known solution is to choose one of the elements of $\boldsymbol{b}$, say $b_3$, as the reference point and fix $b_3 = 1$.  Denote the constrained version of the model by
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
X_{t, k}(1) &=& \gamma_k(1) X_{t-1, k}(1) + w_{t, k}\\
\boldsymbol{v}_{t, k} | \sigma^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k(1) \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}\left(0, \tau^2_k(1)\right),
\end{eqnarray*}
where the trailing input parameter emphasizes the constraint $b_{3} = 1$. Since this version is identifiable, estimates of the model parameters can be obtained. Denote the estimates by placing a hat on the parameter symbol. For instance, $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ represent the estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$, respectively. These estimates are found via Gibbs sampling that only makes use of standard distributions. See Appendix \ref{appendix} for the technical details of the sampling step, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.

\subsection{Calibration Step}
\label{calibration_step}
Since the model parameters can be estimated by fixing $b_3$ to any constant, the next step is to search for the constant that gives an optimally sharp and calibrated estimate of the hidden process.  This section introduces an efficient procedure that finds the optimal constraint without requiring any additional runs of the sampling step. First, assume that parameter estimates  $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ have already been obtained via the constrained sampling step described in Section \ref{sampling_step}. Since for any $\beta \in \R / \{0\}$,
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
 &=&   \boldsymbol{M}_k \left(\boldsymbol{b} (1) \beta \right) \left(X_{t, k}(1)/\beta\right)  + \boldsymbol{v}_{t, k}\\
&=&   \boldsymbol{M}_k \boldsymbol{b} (\beta) X_{t, k}(\beta) + \boldsymbol{v}_{t, k},
\end{eqnarray*}
the parameter values under $b_{3} = \beta$ can be obtained from $\boldsymbol{b} (\beta) = \boldsymbol{b} (1) \beta $ and $X_{t, k}(\beta) = X_{t, k}(1)/\beta$. 
%Furthermore, an estimate of $X_{t, k}(\beta)$ is given by $\hat{X}_{t, k}(1)/\beta$. 
This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$. Since the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated with the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$. A natural criterion for this maximization is given by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}). Due to the possibility of \textit{complete separation} in any one question (see, e.g., \citet{gelman2008weakly}), the maximization must be performed over multiple questions. Therefore,
\begin{eqnarray}
 \hat{\beta}&=&  \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / \beta \right) \label{OSE}
\end{eqnarray}
where $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. The function $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
\begin{eqnarray*}
S_{BRI}(Z, X) &=& -(Z - \logit^{-1}(X))^2
\end{eqnarray*}
or the logarithmic score (\citet{good1952rational})
\begin{eqnarray*}
S_{LOG}(Z, X) &=& Z \log\left(\logit^{-1}(X)\right) + (1-Z) \log\left(1-\logit^{-1}(X)\right)
\end{eqnarray*}
Since it is not clear which rule should be used for predicting geopolitical events, the \textit{Sample-Then-Calibrate} procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}. Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
\begin{eqnarray}
 \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{\beta} \nonumber\\
 \hat{\boldsymbol{b}}_{t,k}&=& \hat{\boldsymbol{b}}(1) \hat{\beta} \nonumber\\
   \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1) \hat{\beta}^2\nonumber\\
  \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\nonumber\\
  \hat{\gamma}_{k}&=& \hat{\gamma}_{k}(1)\nonumber
\end{eqnarray}
Notice that estimates of $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint. Therefore their constrained and unconstrained versions are the same.
